---
title: "Saving Visualizations"
output: html_document
---

<br>

```{r, message = F}
library('tidyverse')
```

<br>

These are base R, except the ggsave, which is ggplot2. To show, we'll create a sample plot. Why not just use the GUI? you can easily hit the export button in the RStudio GUI, but, for reproducibility, you may want to use the code. 

<br> 

```{r}
set.seed(1)
x <- 1:100
y <- 1 + 2*x + rnorm(100,0,2)
```

<br>

### Saving Images in R 

<br> 


### using ggsave

Why use ggsave? easier syntax, more straightfoward a process. no need to turn device on and off. 

```{r, eval = F}
ggplot() + 
  geom_point(aes(x,y))

ggsave('filename.png')
```

<br> 

Best thing about ggsave is that it has sensible defaults. If you want to specify the arguements, here are the useful ones. 

<br> 

```{r, eval = F}
ggsave('filename.png',       # name of file
       plot = last_plot(),   # this is the default
       device = png(),       # png, jpeg, pdf, etc
       height = 6.67,        # specify height
       width = 6.67,         # specify width
       units = 'in',         # specify units, uses default
       scale = 1             # specify a scaling vector 
       )
```

<br> 

### using pdf

```{r, eval = F}
# first declare the pdf function
pdf('filename.pdf', # saved files name
    7,              # width of image in inches
    5               # height of image in inches
    )

# next plot the visualization
ggplot() + 
  geom_point(aes(x,y))

# lastly, turn off the pdf 'device'
dev.off()
```

<br>

The .pdf has been saved to my working directory. I could have specified a path in the filename. 

<br> 

### using jpg, png, or bmp

```{r, eval = F}
# first specify the parameters of your image
jpeg('filename.jpg',  # file name and path
     width = 480,     # width in units
     height = 480,    # height in units
     units = 'px',    # units can be pixels, inches, cm, or mm
     quality = 0.9,   # quality of compresson as percentage
)

# plot the visualization
plot(x,y)

# lastly, turn off the jpg device
```

all use the same syntax, just change the 'jpeg' to 'png' or 'bmp'. 
