---
title: 'Data Structures'
output: html_document
---

A bit more confusing than the atomic vectors, data structures are built on top of atomic vectors. can be categorized into homogenous (containing all of the same atomic vector) and heterogeneous (containing different types of atomic vectors). 

First we'll look at homogeneous. 

<br>

## Homogeneous

<br>

### Matix

Just as you'd expect from linear algebra, matrices. 

```{r}
matrix_dbl <- matrix(c(1, 1, 1, 0), nrow = 1)
matrix_int <- matrix(c(1L,2L, 3L, 4L, nrow = 4))
matrix_chr <- matrix(c('a', 'a', 'a', 'a'), ncol = 2)
matrix_log <- matrix(c(T, T, T, F), nrow = 2, ncol = 2)
```

<br>

If you try to make it homogenous, coercion rules apply, see atomic vector note. 

```{r}
coerced_matrix <- matrix(c(1, T, 'Seven', 5L), nrow = 1)
coerced_matrix
```

<br>

Attributes are different depending on the function you apply to it.

Is both a matrix and numeric.

```{r}
is.matrix(matrix_dbl)
is.numeric(matrix_dbl)
```

<br>

But this type of coercion wont work. 

```{r, error = T}
matrix_dbl == matrix_log
```

<br>

But this will

```{r}
sum(matrix_log) == sum(matrix_dbl)
```

<br>

```{r}
dim(matrix_chr)
```

<br>

```{r}
matrix_int[4,1]
```

<br>


### Array

<br>

To me, these are the most confusing data structure in R. A multi dimensional matrix. basically stacked matrices. 

<br>

Notices homogeneous

```{r}
coerced_array <- array(c(matrix_dbl, matrix_int, matrix_log, matrix_chr))
is.array(coerced_array)
typeof(coerced_array)
coerced_array
```

<br>

Multi-layer
```{r}
multi_layer <- array(rep(matrix_dbl, 4), dim = c(1,4,3))
multi_layer
```

<br>

pulling the first row, fourth column, 3rd 'layer'
```{r}
multi_layer[1,4,3]
```

<br>

as you can imagine, these get complex.

<br>

## Hetrogeneous

<br>

### List

<br> 

a big of an ugly catch all, retains structures and types

All to a single element in the list
```{r}
single_list <- list(c(matrix_dbl, matrix_int, matrix_log, matrix_chr))
is.list(single_list)
str(single_list)
single_list
```

<br>

```{r}
multi_list <- list(matrix_dbl, matrix_int, matrix_log, matrix_chr)
is.list(multi_list)
str(multi_list)
multi_list
```

<br>

### Difference between [[]] and []

when using a list and other things, you might use [[]] or []. [[]] pulls the list itself, [] shows you whats inside. 

see how [[]] works with a second set of indicies because now you're working with that specific matrix, while [] does not
```{r, error = T}
multi_list[[1]][1,2]
multi_list[1][1,2]
```

<br>

### Data Frame

most common data structure. 

```{r}
df <- data.frame(a = 1:3,
                 b = c('chr_1', 'chr_2', 'chr_3'),
                 c = c(3L, 2L, 1L),
                 d = c(T,F,T))
df
```

<br>

renaming

```{r}
colnames(df) <- c('name_1', 'name_2', 'name_3', 'name_4')
df
```

not advised to name the rows. use the indices as unique identifiers, if they have names, make it a variable. 

<br>

combining dataframes

column bind

```{r}
new_df <- data.frame(name_5 = c(5,5,5))

df <- cbind(df, new_df)
df
```

row bind - have more observations?
```{r}
new_row <- data.frame(name_1 = 4, 
                      name_2 = 'chr_4', 
                      name_3 = 0L, 
                      name_4 = F, 
                      name_5 = 5
                      )

df <- rbind(df, new_row)
df
```

access data frame 
```{r}
df[1]
df[[1]]
df$name_1
```

<br>

Alternatively, use tibbles. see that note. 