---
title: 'Gather and Spread'
output: html_document
---

### `gather()`

Often, a dataset has column names that are not the names of variables, but the values of variables. take `table4a` for example. 

```{r}
table4a
```

<br> 

In `table4a`, the column names `1999` and `2000` are values of the variable year, not variables themselves. The `gather()` function can fix this, it is perhaps the most commonly used function in tidyr. Begin by listing the column names that are actually variable names into `gather()` as its first arguments. Once these column names have been entered, define a `key` as the second argument. In this case, `1999` and `2000` are both years, so the key, or new variable name, is `year`. Finally, define a `value` to be associated with the key as the last argument in `gather()`. The `value` is the name given to the elements that were previously linked to the previous `1999` and `2000` false variables. Because these two years have been gathered into a new `year` variable, the associated elements need a new column to call home. In this case, the values represent a number positive cases, so the `value` is now defined as cases. 

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```

<br  />

### `spread()`


When an observation is scattered across multiple rows, variable names appear repeatedly as elements in a dataframe. In `table2` a single observation is a country in each year, but each observation is spread across two rows. 

```{r}
table2
```

<br  />

To tidy this data, utilize `spread()`, one of the most common functions in tidyr. The first argument in `spread()` is the column that contains the variable names, the `key` column. In the case of `table2`, the key column is type, because cases and population are both variables, not values. The second and final argument in `spread()` is the column that contains values associated with the previously defined key, the `value` column. In `table2` the it is the count column that contains the associated values.

```{r}
table2 %>%
spread(key = type, value = count)
```

<br  />

