---
title: 'Missing Values'
output: html_document
---

### `fill()`

When dealing with missing data, it can be the case that you know that missing values are supposed to be carried on from the last observation. Something along the line of "ditto" quotations on a sign-up sheet. In the tibble treatment, we see just that. 

```{r, echo = FALSE}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
```

```{r}
treatment
```


The function `fill()` is the perfect fix for this situation. `fill()` takes a set of columns where you want missing values to be replaced with the most recent non-missing value. Simply input the column in question as the argument in `fill()`, and let R do the rest. In the case of the tibble `treatment`, the column in question is `person`. 

```{r}
treatment %>%
  fill(person)
```


<br>

### `complete()`

When dealing with missing data it's often important to turn implicitly missing values to explicit missing values. There are two missing values from the stocks tibble, 4th quarter 2015 and 1st quarter 2016. 

```{r, echo = FALSE}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
````

```{r}
stocks
```

<br> 

The `complete()` function takes a set of columns, and finds all unique combinations. It ensures the original dataset contains all those values, explicitly filling in `NA` when necessary. The input arguments of `complete()` are simply the columns you want to cross reference. In the case of `stocks` we want to find all of the combinations between the `year` and `qtr` variable, as to fill in implicit missing variables with `NA`. 

```{r}
stocks %>% 
  complete(year, qtr)
```


