---
title: 'Separate and Unite'
output: html_document
---

<br>

```{r}
library('tidyverse')
```

### separate

Sometimes a column contains more than variable worth of information in each element. In `table3` this is exemplified by the `rate` column. 

```{r}
table3
```

<br  />

A simple problem to fix, the `separate()` function splits two variables apart from a single column. The first argument in `separate()` is the column in question, in this case it's `rate`. The next argument is a bit more complex; the `into` argument needs a concatenated input, `c()` to define the names and number of new columns. As for table3, the new columns are defined as `into = c("cases", "population")`. Finally, a separator character can be defined in the third argument. The separator defaults to any non-alphanumeric character, but can be customized. 

```{r, warning=FALSE}
table3 %>%
  separate(rate, into = c("cases", "population"))
```

<br  />

### unite

When a single variable is spread between multiple columns, it's time to unite them. In `table5` the century and year variables should be united to create a single variable, the full year. 


```{r}
table5
```

<br  /> 

The `unite()` function works in a similar way to `separate()`, as it is it's inverse. The first argument of `unite()` is to define the name of the new, united, variable. In the case of `table3`, this variable shall be called `full_year`. The next arguments define the columns that need to be united; this can be any number of columns. `century` and `year` need to be united to form `full_year` in this example. Finally, it's sometimes wise to indicate a separator character, just as with `separate()`. The separator character in `unite()` defaults to the underscore, `_`. 

```{r}
table5 %>%
  unite(full_year, century, year, sep = "")
```

<br  /> 