---
title: "advanced joins"
output: html_document
---

<br>

```{r, message = F}
library('tidyverse')
library('dplyr')
```

<br>

```{r}
tib_1 <- tibble(x = 1:2, y = 2:1)
tib_1
tib_2 <- tibble(x = c(1,3), a = 10, b = "a")
tib_2
```

<br>


#### Full Join

The most inclusive join of all, `full_join()` joins every observation from both tables, as the name suggests. 

```{r}
full_join(tib_1, tib_2, by = "x")
```

<br> 

#### Semi Join

The first of two filter joins, `semi_join()`, keeps all observations in the `x` argument that have a match in the `y` argument. In this example, the key is the "x" column of `tib_1`; the only row that matches in `tib_2` is the observation x = 1. Thus, the resulting semi_join only has one observation filtered from `tib_1`. 

```{r}
semi_join(tib_1, tib_2, by = "x")
```

<br> 

#### Anti Join

`anti_join()` is the antithesis of `semi_join()`, the function only returns observations that are not present in the `y` dataset. 

```{r}
anti_join(tib_1, tib_2)
```

<br> 