---
title: 'dplyr group by'
output: html_document
---

<br>

```{r, message = F}
library('tidyverse')
library('nycflights13')
```

<br>


#### Group By

`summarize()` becomes extremely useful when paired with the final primary dplyr function, `group_by()`. `group_by()` changes the unit of analysis from the complete dataset to an individual group, thus changing the scope of summarize. In the above example, the flights dataset is grouped by the `dest` variable, destination. Grouping by destination by itself does nothing to the data frame, so the then pipe, %>%, is used to push the output to the `summarize()` function. Summarize first creates a count variable that is equivalent to the function `n()`. `n()` is another one of those secondary dplyr functions that often comes in handy; `n()` is a function that finds the number of observations in the current group. Next, `summarize()` creates a variable named distance based off the mean distance travelled by each observation, as grouped by destination. Finally, `summarize()` creates a variable named delay, based off the mean arrival delay each observation experienced, as grouped by destination. The resulting data frame gives excellent insight into each of the 105 destinations present in the flights dataset. 


```{r}
group_by(flights, dest) %>%                       
  summarize(count = n(),                          
            dist = mean(distance, na.rm = TRUE),  
            delay = mean(arr_delay, na.rm = TRUE) 
            ) 
```

<br  /> 

